---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  fig.width = 12
)
```

# myspotify

<!-- badges: start -->

[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental) [![CRAN status](https://www.r-pkg.org/badges/version/myspotify)](https://CRAN.R-project.org/package=myspotify)
[![R-CMD-check](https://github.com/audreykim640/myspotify/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/audreykim640/myspotify/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of `myspotify` is to to read personal Spotify data and complete a series of analyses. Spotify is a popular music and podcast streaming app globally.

## What's this package for?

Spotify, a popular music streaming platform, offers users the option to download their personal usage history data. `myspotify` seeks to provide tools to allow users to read and visualize their data to learn more about their listening habits. The idea is somewhat similar to "Spotify Wrapped," a personalized summary that Spotify releases annually, but this can assess history beyond just a year and allows the user to delve into the patterns of the songs they listen to on Spotify.

## Who should use this package?

Anyone who uses Spotify, wants to learn more about their listening habits, and has a basic background in (or is willing to learn!) R.

“Basic background” might include understanding variable assignment (`<-`) and various Tidyverse packages like dplyr and ggplot2. Tidyverse familiarity is not necessary but would allow for more data exploration opportunities.

## How to install this package?

```{r loading packages, include=FALSE}
library(devtools)
devtools::install_github("audreykim640/myspotify")
library(dplyr)
library(ggplot2)
library(myspotify)
```


You can install the development version of myspotify from [GitHub](https://github.com/) with:

``` {r}
# install.packages("devtools")
devtools::install_github("audreykim640/myspotify")
```

## What datasets are there?

-   `extended_example`: example dataset of Spotify listening history from one of the authors
-   Load your own!
    -   After loading the package and downloading your data from Spotify (instructions [here](https://github.com/audreykim640/myspotify/blob/main/vignettes/myspotify.Rmd)), run your data through the `read_file()` function:

```{r example_data_download}
# example:
# my_data <- read_file("workspace/my_spotify_data/myData/endsong_0.json",
#                      "workspace/my_spotify_data/myData/endsong_1.json",
#                      "workspace/my_spotify_data/myData/endsong_2.json")
```

**Other** `myspotify` **functions will not run without this step**. The assigned variable (`my_data` in the example above) is what you can use in conjunction with all of the other functions in the package.

## Example data visualizations

```{r examples_pt1}
artist_trends(extended_example)

# Spotify data by default includes all songs that started playing
# Say we only want to see artists whose songs we played for at least 2 minutes...
extended_example %>% 
  filter_playtime(length = 1) %>% 
  artist_trends()
```

```{r examples_pt2}
# Plot the top 5 most skipped songs
skipped_songs <- extended_example %>% 
  most_skipped() %>% 
  arrange(desc(skips)) %>% 
  head(5)
  
ggplot(skipped_songs, aes(x = song, y = skips)) +
  geom_col(fill= "darkmagenta") +
  theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
  theme(axis.title = element_text(face="bold"))
```

```{r examples_pt3} 
#Search for a specific word or phrase in a song name 
#Returns a data frame

search_word(extended_example, "Christmas")

# Returns all songs streamed with "Christmas" in the title
```

```{r examples_pt4}
# Make a plot of the top ten artists/bands you have streamed of all time

most_played_artists(extended_example)
```


## Note about testing

There's a test for the dataset in this package, but the rest of the package (functions) return plots. We did not write tests for these, as the suggested package, `vdiff`, does not exist for our working version of R.
